---
title: "Retirement Income Toolkit - Economic Scenario - VAR"
author: "Queenie Huang"
output:
 bookdown::html_document2:
    toc: yes
    toc_depth: '3'
    number_sections: yes
    fig_caption: yes
    code_folding: hide
urlcolor: blue
header-includes: |    
 \usepackage{amsmath,amsthm,amssymb,amsfonts, mathtools, empheq, bbm}
 \newcommand{\vect}[1]{\boldsymbol{#1}}
---

<style type="text/css">

body{ /* Normal  */
      font-size: 15px;
  }
td {  /* Table  */
  font-size: 15px;
}
h1.title {
  font-size: 30px;
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 15px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 15px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(scipen = 1, digits = 6) # use 2 d.p. in output
```

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
library(vars)
library(tseries)
library(corrplot)
library(kableExtra)
library(aTSA)
```

A VAR(p) process can be represented by \begin{align*}
    \vect Y_t = \vect \alpha + A_1 \vect Y_{t-1} + \ldots + A_p \vect Y_{t-p} + \vect \epsilon, 
\end{align*} where $n$ is the number of economic/financial variables, $p$ is the lag, $\vect \epsilon$ is a white noise process, $A_1, \ldots, A_p$ are fixed coefficient matrices. 

# Economic/financial variables 

#

