---
title: "Retirement Income Toolkit - Economic Scenario - Pricing Kernels"
author: "Queenie Huang"
output:
 bookdown::html_document2:
    toc: yes
    toc_depth: '4'
    number_sections: yes
    fig_caption: yes
    code_folding: hide
urlcolor: blue
header-includes: |    
 \usepackage{amsmath,amsthm,amssymb,amsfonts, mathtools, empheq, bbm}
 \newcommand{\vect}[1]{\boldsymbol{#1}}
---

<style type="text/css">

body{ /* Normal  */
      font-size: 15px;
  }
td {  /* Table  */
  font-size: 15px;
}
h1.title {
  font-size: 30px;
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 15px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 15px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(scipen = 1, digits = 6) # use 2 d.p. in output
```

```{r}
library(tidyverse)
library(readxl)
library(zoo)
library(janitor)
library(doParallel)
e1 = as.matrix(c(1, rep(0,7)))
```


# Import Data 

```{r}
start_pc_date = as.Date("1993-04-01")
end_date = as.Date("2021-03-31")
max_maturity = 40
n_index = c(1,4,8,20,40)
maturities = c("Date", paste(1:max_maturity, "qtrs", sep = "_"))
var_names = c("zcp3m_yield", "zcp10y_spread", "home_index_growth", "rental_yield", "GDP_growth", "CPI_growth", "ASX200_growth", "AUD_growth")


#######################
# import interest rates 
zcp_old = read_excel("ZCP_till2008.xls")[-c(1:10), -2]
zcp_new = read_excel("ZCP_from2008.xls")[-c(1:10), -2]
names(zcp_old) = maturities
names(zcp_new) = maturities
zcp = rbind(zcp_old, zcp_new)
rm(zcp_old, zcp_new)

# cleaning 
zcp = zcp %>% mutate_if(is.character, as.numeric)
zcp$Date = as.Date(zcp$Date, origin = "1899-12-30")
zcp = clean_names(zcp)

# match the time window 
zcp = zcp %>% filter(date <= end_date & date >= start_pc_date)
zcp = zcp %>% mutate(Qtr = as.yearqtr(date)) %>% group_by(Qtr) %>% summarise_at(vars(-date), mean) 

# add row names & change class()
time_names_past = zcp$Qtr
zcp_sdf = as.matrix(zcp[-1,n_index+1]) / 100
zcp = as.matrix(zcp[-1,-1]) / 100
row.names(zcp) = as.character(time_names_past[-1])
row.names(zcp_sdf) = as.character(time_names_past[-1])

###################
# import zt vectors 
all_data = read.csv("data/all_data.csv")[,-1]
all_data = as.matrix(t(all_data))
row.names(all_data) = var_names
colnames(all_data) = as.character(time_names_past)
all_data[1,] = all_data[1,] / 100
all_data[2,] = all_data[2,] / 100

##############
# VAR Dynamics
# intercept 
intercept = matrix(c(-0.001656651,0.000142228,0.000697426,0.001208242,0.007404573,0.006936114,0.020845829, -0.001948064), nrow = 8, ncol = 1)
# coefficient matrix in VAR(1)
coef = data.frame(zcp3m_yield = c(0.339929,0.1137847,0.02865682, -0.1962058, -0.0213723,0.00446621,0.00991223,0.00654165),
                  zcp10y_spread = c(-0.451455,0.8670374,0.02193637,0.0199211,0.0442296,0.05746370,-0.00562332,0.01852652),
                  home_index = c(-0.661822,0.0568036,0.72566350,0.0746341,0.3094571, -0.17668326,0.03114352,0.00984050),
                  rental_yield = c(0.150427, -0.1065600, -0.00245535,0.1967380, -0.0335203,0.04828259,0.00803609,0.00148116),
                  GDP = c(-0.243469,0.1739440, -0.06584194,0.3091203, -0.0617071, -0.15061791,0.01758549,0.04417218),
                  CPI = c(0.256898, -0.0754316,0.00615750, -0.1858264, -0.0911666,0.08844679,0.00387475, -0.00201055),
                  ASX200 = c(-0.167551,0.1251294, -0.32533013, -1.1531345,0.1320861, -1.93119670,0.26973160,0.12588817),
                  AUD = c(-2.216189,0.3490000,  0.26377394, -1.1205875,0.2365210, -0.77832396, -0.01069297,0.24006978),
                  row.names = var_names)
colnames(coef) = var_names
coef = as.matrix(t(coef))

# coefficient matrix in VAR(2)
coef1 = matrix(0, nrow = 8, ncol = 8)
coef1[1,1] = 1
coef1[4,4] = 1
coef1 = coef + coef1
coef2 = matrix(0, nrow = 8, ncol = 8) 
coef2[,1] = -coef[,1]
coef2[,4] = -coef[,4]

# residual covariance matrix 
covres = data.frame(zcp3m_yield = c(1.47783e-05,-1.01017e-05,2.53209e-06,-2.71426e-06,6.60361e-06,4.41163e-06,0.0000432530,7.57409e-05),
                    zcp10y_spread = c(-1.01017e-05,2.11857e-05,-6.23599e-07,-1.20674e-06, -4.37033e-06, -1.44520e-06, -0.0000266498, -4.78746e-05),
                    home_index = c(2.53209e-06,-6.23599e-07,1.43124e-04,-6.82911e-06,8.61086e-06, -1.00958e-06,0.0001237822,4.30438e-05),
                    rental_yield = c(-2.71426e-06,-1.20674e-06,-6.82911e-06,1.67679e-05, -3.35037e-08,  2.46972e-06, -0.0000480865, -2.81226e-05),
                    GDP = c(6.60361e-06,-4.37033e-06,8.61086e-06,-3.35037e-08,9.90145e-05,2.00431e-05,0.0001435020,7.72833e-06),
                    CPI = c(4.41163e-06,-1.44520e-06,-1.00958e-06,2.46972e-06,2.00431e-05,3.09181e-05,0.0000535892,2.50834e-05),
                    ASX200 = c(4.32530e-05,-2.66498e-05,1.23782e-04,-4.80865e-05,1.43502e-04,5.35892e-05,0.0029571153,9.07070e-04),
                    AUD = c(7.57409e-05,-4.78746e-05,4.30438e-05,-2.81226e-05,7.72833e-06,2.50834e-05,0.0009070698,1.45155e-03),
                    row.names = var_names)
colnames(covres) = var_names
covres = as.matrix(t(covres))
```

# Model - estimate lambda_0, lambda_1 (historical data)

```{r}
######################
# calculate An, Bn, Cn

get_ABC = function (lambda_0, lambda_1) {
  lambda_0 = as.matrix(lambda_0)
  lambda_1 = as.matrix(lambda_1)
  
  # initialise dataframes to store An, Bn, Cn
  An = matrix(NA, nrow = 1, ncol = max_maturity)
  Bn = matrix(NA, nrow = length(intercept), ncol = max_maturity)
  Cn = matrix(NA, nrow = length(intercept), ncol = max_maturity)
  
  # initial values A1, B1, C1
  An[1] = 0
  Bn[,1] = -e1
  Cn[,1] = rep(0,length(intercept))
  
  # terminal values A40, B40, C40
  # An[max_maturity] = 0
  # Bn[,max_maturity] = -max_maturity * e1 - max_maturity * c(0,1,rep(0,length(intercept)-2))
  # Cn[,max_maturity] = rep(0,length(intercept))
  # 
  # calculate recursively
  for (n in 1:(max_maturity-1)) {
    An[n+1] = An[n] + t(Bn[,n]) %*% (intercept - chol(covres) %*% lambda_0) + 1/2 * t(Bn[,n]) %*% covres %*% Bn[,n]
    Bn[,n+1] = -e1 + t(coef1 - chol(covres) %*% lambda_1) %*% Bn[,n] + Cn[,n]
    Cn[,n+1] = t(coef2) %*% Bn[,n]
  }
  
  # output 
  output = list(An, Bn, Cn)
  names(output) = c("An", "Bn", "Cn")
  output = lapply(output, function(x){x = as.matrix(x)})
  output = lapply(output, function(x){colnames(x) = colnames(zcp); return (x)})
  return (output)
}

#############################################
# calculate sum (r(n)t hat - r(n)t) ^ 2 stats

# only for r(1), r(4), r(8), r(20), r(40)
# get_rnt = function (lambda_0, lambda_1) {
#   # An, Bn, Cn 
#   ABCn = get_ABC(lambda_0, lambda_1)
#   An = ABCn$An
#   Bn = ABCn$Bn
#   Cn = ABCn$Cn
#   
#   # zt: all_data
#   
#   # initialise dataframe to store historical r_t^(n)
#   rnt = matrix(NA, nrow = length(time_names_past), ncol = max_maturity)
#   row.names(rnt) = as.character(time_names_past)
#   colnames(rnt) = maturities[-1]
#   
#   # r(n)t hat  
#   for (t in 2:length(time_names_past)) {
#     for (n in n_index) {
#       rnt[t,n] = -An[n] / n - t(Bn[,n]) %*% all_data[,t] / n - t(Cn[,n]) %*% all_data[,t-1] / n
#     }
#   }
#   
#   return (rnt[-1,n_index])
# }

# r(1), r(2), ..., r(40)
get_rnt = function (lambda_0, lambda_1) {
  # An, Bn, Cn
  ABCn = get_ABC(lambda_0, lambda_1)
  An = ABCn$An
  Bn = ABCn$Bn
  Cn = ABCn$Cn
  
  # zt: all_data
  
  # initialise dataframe to store historical r_t^(n)
  rnt = matrix(NA, nrow = length(time_names_past), ncol = max_maturity)
  row.names(rnt) = as.character(time_names_past)
  colnames(rnt) = maturities[-1]
  
  # r(n)t hat  
  for (t in 2:length(time_names_past)) {
    for (n in 1:max_maturity) {
      rnt[t,n] = -An[n] / n - t(Bn[,n]) %*% all_data[,t] / n - t(Cn[,n]) %*% all_data[,t-1] / n
    }
  }
  
  return (rnt[-1,])
}

get_sse = function (lambda_list) {
  # reorganise parameters
  lambda_0 = matrix(lambda_list[1:8], nrow = 8)
  lambda_1 = matrix(lambda_list[-c(1:8)], nrow = 8, ncol = 8, byrow = T)
  
  rnt_hat = get_rnt(lambda_0, lambda_1)
  sum((rnt_hat - zcp) ^ 2)
}

#################################
# minimise for lambda_0, lambda_1
# lambda_list = list(lambda0_1=0, lambda0_2=0, lambda0_3=0, lambda0_4=0, lambda0_5=0, lambda0_6=0, lambda0_7=0, lambda0_8=0,
#                    lambda1_11=0, lambda1_12=0, lambda1_13=0, lambda1_14=0, lambda1_15=0, lambda1_16=0, lambda1_17=0, lambda1_18=0,
#                    lambda1_21=0, lambda1_22=0, lambda1_23=0, lambda1_24=0,  lambda1_25=0, lambda1_26=0, lambda1_27=0, lambda1_28=0,
#                    lambda1_31=0, lambda1_32=0, lambda1_33=0, lambda1_34=0, lambda1_35=0, lambda1_36=0,  lambda1_37=0, lambda1_38=0,
#                    lambda1_41=0, lambda1_42=0, lambda1_43=0, lambda1_44=0, lambda1_45=0, lambda1_46=0, lambda1_47=0, lambda1_48=0,
#                    lambda1_51=0, lambda1_52=0, lambda1_53=0, lambda1_54=0, lambda1_55=0, lambda1_56=0, lambda1_57=0, lambda1_58=0,
#                    lambda1_61=0, lambda1_62=0, lambda1_63=0, lambda1_64=0, lambda1_65=0, lambda1_66=0, lambda1_67=0, lambda1_68=0,
#                    lambda1_71=0, lambda1_72=0, lambda1_73=0, lambda1_74=0,  lambda1_75=0, lambda1_76=0, lambda1_77=0, lambda1_78=0,
#                    lambda1_81=0, lambda1_82=0, lambda1_83=0, lambda1_84=0, lambda1_85=0, lambda1_86=0,  lambda1_87=0, lambda1_88=0)
# 
# optim_ctrl = list(maxit = 500)
# cl <- makeCluster(detectCores())
# registerDoParallel(cl)
# lambdas = optim(lambda_list, get_sse, control = optim_ctrl)
# stopCluster(cl)
# lambdas
# 
# save(lambdas, file = "data/lambdas.RData")
# load("data/lambdas.RData")
lambda_0 = as.matrix(read.csv("data/lambda_0.csv")[,-1])
lambda_1 = as.matrix(read.csv("data/lambda_1.csv")[,-1])
```

# Model - fitted s_t

```{r}
############
# fitted VAR 
# zt_fitted = matrix(NA, nrow = 8, ncol = 112)
# zt_fitted[,1:2] = all_data[,1:2]
# zt_fitted[1:2,1:2] = all_data[1:2, 1:2] * 100
# for (i in 3:112) {
#   zt_fitted[,i] = intercept + coef1 %*% zt_fitted[,i-1] + coef2 %*% zt_fitted[,i-2]
# }
zt_fitted = read.csv("data/all_data_fitted.csv")[,-1]
row.names(zt_fitted) = var_names
colnames(zt_fitted) = time_names_past
zt_fitted[1:2,] = zt_fitted[1:2,] / 100

###########
# fitted st
lambda_t = sapply(1:112, function (x) {lambda_0 + lambda_1 %*% zt_fitted[,x]}, simplify = T)
st_fit = sapply(1:112, function (x) {exp(-zt_fitted[1,x] - 1/2 * sum(lambda_t[,x]^2))})
plot(time_names_past, st_fit * 100, "l", main = "Fitted SDFs", xlab = "", ylab = "")
```

# Model - historical s_t

```{r}
lambda_t = sapply(1:112, function (x) {lambda_0 + lambda_1 %*% all_data[,x]}, simplify = T)
st_past = sapply(1:112, function (x) {exp(-all_data[1,x] - 1/2 * sum(lambda_t[,x]^2))})

##############
# correlations
# cor(st_past, all_data[1,])
# cor(st_past, all_data[2,])
# cor(st_past, all_data[3,])
# cor(st_past, all_data[4,])
# cor(st_past, all_data[5,])
# cor(st_past, all_data[6,])
# cor(st_past, all_data[7,])
# cor(st_past, all_data[8,])
```


# Testing: get_var_simulations

```{r}
num_paths = 10
num_years = 20

source("../SDF_Simulation.R")
source("../VAR.R")

cl <- makeCluster(detectCores())
registerDoParallel(cl)
set.seed(1)
sdf = get_var_simulations(num_paths = num_paths, num_years = num_years, return_sdf = T)
stopCluster(cl)

par(mfrow = c(2,2))

#######
# zcp3m
plot_date = c(as.Date(time_names_past), row.names(sdf$zcp3m_yield)[-1])
plot_data = c(all_data[1,], sdf$zcp3m_yield[-1,1])
plot(plot_date, plot_data, "l", main = "zcp3m", col = "grey")
for (i in 2:num_paths) {
  plot_data = c(all_data[1,], sdf$zcp3m_yield[-1,i])
  lines(plot_date, plot_data, col = "grey")
}

########
# zcp10y
plot_data = c(all_data[2,], sdf$zcp10y_spread[-1,1])
plot(plot_date, plot_data, "l", main = "zcp10y", col = "grey")
for (i in 2:num_paths) {
  plot_data = c(all_data[2,], sdf$zcp10y_spread[-1,i])
  lines(plot_date, plot_data, col = "grey")
}

######
# rent 
plot_date = c(as.Date(time_names_past), row.names(sdf$rental_yield)[-1])
plot_data = c(all_data[4,], sdf$rental_yield[-1,1])
plot(plot_date, plot_data, "l", main = "rent", col = "grey")
for (i in 2:num_paths) {
  plot_data = c(all_data[4,], sdf$rental_yield[-1,i])
  lines(plot_date, plot_data, col = "grey")
}

#####
# sdf 
plot_data = c(st_past, sdf$discount_factors[-1,i])
plot(plot_date, plot_data, "l", main = "SDFs", xlab = "", ylab = "", col = "grey")
for (i in 2:num_paths) {
  plot_data = c(st_past, sdf$discount_factors[-1,i])
  lines(plot_date, plot_data, col = "grey")
}
```
